### 2019-09-30 ThesisJo/2_background.tex
2019-09-30

A few solutions to the wave equation is of interest, in particular that of a plane and spherical wave,
%
\begin{align}
f_\textrm{plane}(\vec{p},t)  &= A e^{j(\omega t - \vec{k} \cdot \vec{p})} \\
f_\textrm{spheric}(\vec{p},t) &= \frac{A}{\lVert \vec{p} \rVert} e^{j(\omega t - \lVert\vec{k}\rVert \lVert \vec{p} \rVert)},
\end{align}
% 
where $A$ is amplitude, $\omega$ is temporal wave frequency, $t$ is time, $\lVert\vec{k}\rVert=\frac{\omega}{c}=\frac{2\pi}{\lambda}$ is spatial wave frequency (wavenumber) and $\vec{p}$ is position of the source relative to the observer.

@@

The beamformer output $z[n]\in\mathbb{C}$ is by definition the weighted sum of all the delayed data samples:
%
\begin{align}
z[n] = \dvec{w}^H[n]\dvec{x}[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n]\\x_1[n]\\\vdots\\x_{M-1}[n]},\label{2_eq_z}
\end{align}
%
where $w_m\in\mathbb{C}$ is the weight factor assigned to channel $m$. With static weights this would be referred to as the conventional delay-and-sum (DAS) beamformer. A large variety of weighting functions exist for trading lateral resolution for improved noise suppression (contrast), but a compromise between the two is unavoidable~\cite{Harris1978}. However, there exists a class of adaptive beamformers that target this limitation by adapting the weights to the impinging wavefield. The MVDR method is one such technique.

@@

\begin{figure}[t]\centering%
\graphicsAI[drawing,width=0.7\linewidth]{gfx/implementation.svg}%
\caption{MVDR beamforming. For one of the total number of pixels in range and azimuth, $N_y$ and $N_x$,\newline
1. an $L\times{}L$ sample covariance matrix $\eR$ is computed, \newline
2. the term $\eR^{-1}\1$ is found using a linear equation solver,\newline
3. and the beamformer output is computed from $z$ from (\ref{finalZ}), where $\w$ is found by substituting $\eR^{-1}\1$ into (\ref{2_eq_weights}). }\label{mvdr_beamforming}
\end{figure}
%
The basic idea behind the MVDR method is to compute the set of weights that minimizes noise and signal power accumulated by the array, while ensuring unity gain for signals coming from the pixel of interest. When the data is delayed this can be formulated as follows:
%
\begin{gather}
\argmin{\dvec{w}}\ E\Big\{\big|\dvec{w}^H[n]\dvec{x}[n]\big|^2\Big\} \qquad\text{subject to}\qquad \dvec{w}^H[n]\dvec{1} = 1. \label{2_eq_mvdr_def}
% \vec w[n] = \frac{\Ri[n]\1}{\1^T\Ri[n]\1},\label{2_eq_weights}
\end{gather}
%
This is a convex optimization problem with the solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\dvec{1}}{\dvec{1}^T\Ri[n]\dvec{1}},\label{2_eq_weights}
\end{gather}
%
where $\1$ is a row vector of ones that represents broadside steering, and $\R=E\{\x[n]\x^H[n]\}$ is the spatial covariance matrix for the full array. Since $\R$ is unknown, we estimate it by by computing a sample covariance matrix $\eR$. In this computation we perform some degree of \emph{spatial averaging} to avoid signal cancellation, \emph{temporal averaging} to maintain true speckle statistics, and \emph{diagonal loading} to improve robustness to parameter errors ~\cite{Synnevag2009a}. Combined, these techniques will also ensure a numerically well conditioned $\eR$.

>>>>

A few solutions to the wave equation are of interest, in particular that of a plane and spherical wave:
%
\begin{align}
f_\textrm{plane}(\vec{p},t)  &= A e^{j(\omega t - \vec{k} \cdot \vec{p})} \\
f_\textrm{spheric}(\vec{p},t) &= \frac{A}{\lVert \vec{p} \rVert} e^{j(\omega t - \lVert\vec{k}\rVert \lVert \vec{p} \rVert)},
\end{align}
% 
Here $A$ is amplitude, $\omega$ is temporal wave frequency, $t$ is time, $\lVert\vec{k}\rVert=\frac{\omega}{c}=\frac{2\pi}{\lambda}$ is spatial wave frequency (wavenumber) and $\vec{p}$ is position of the source relative to the observer.

@@

The beamformer output $z[n]\in\mathbb{C}$ is by definition the weighted sum of all the delayed data samples:
%
\begin{align}
z[n] = \dvec{w}^H[n]\dvec{x}[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n]\\x_1[n]\\\vdots\\x_{M-1}[n]} \in\mathbb{C},\label{2_eq_z}
\end{align}
%
where $w_m\in\mathbb{C}$ is the weight factor assigned to channel $m$. With static weights this is referred to as the conventional delay-and-sum (DAS) beamformer. A variety of weighting functions exist for trading lateral resolution for improved noise suppression (contrast), but a compromise between the two is unavoidable~\cite{Harris1978}. A class of adaptive beamformers target this limitation by adapting the weights to the impinging wavefield. One such technique is the minimum variace distortionless response (MVDR) method.

\begin{figure}[tp]\label{II_fig_article_relation}
\makebox[\linewidth][c]{%
\graphicsAI[drawing,width=1.3\linewidth]{gfx/article_relation.svg}}
\caption{Method relations. This thesis explores three beamformers reconstructing images from acoustic echoes: time-domain MVDR (reference), frequency-domain MVDR (close approximation) and time-domain LCA (educated trial-and-error). Background on synthesis of image templates from the reconstructed images will be presented in article IV. }
\end{figure}

% Accelerated adaptivity is key: it enables rapid dynamic responses to mission events. }

@@

\begin{figure}[t]\centering%
\graphicsAI[drawing,width=1\linewidth]{gfx/implementation.svg}%
\caption{MVDR beamforming. For one of the total number of pixels in range and azimuth, $N_y$ and $N_x$,\newline
1. an $L\times{}L$ sample covariance matrix $\eR$ is computed, \newline
2. the term $\eR^{-1}\1$ is found using a linear equation solver,\newline
3. and the beamformer output is computed from $z$ from (\ref{finalZ}), where $\w$ is found by substituting $\eR^{-1}\1$ into (\ref{2_eq_weights}). }\label{mvdr_beamforming}
\end{figure}
%
The basic idea behind MVDR is to compute the set of weights that minimizes noise and signal power accumulated by the array, while seeking unit gain for signals from the pixel of interest. With pre-delayed data this can be formulated as follows:
%
\begin{gather}
\argmin{\dvec{w}}\ E\Big\{\big|\dvec{w}^H[n]\dvec{x}[n]\big|^2\Big\} \qquad\text{subject to}\qquad \dvec{w}^H[n]\dvec{1} = 1. \label{2_eq_mvdr_def}
% \vec w[n] = \frac{\Ri[n]\1}{\1^T\Ri[n]\1},\label{2_eq_weights}
\end{gather}
%
This optimization problem has the solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\dvec{1}}{\dvec{1}^T\Ri[n]\dvec{1}},\label{2_eq_weights}
\end{gather}
%
where $\dvec1$ is a row vector of ones representing broadside steering and $\R=E\{\x[n]\x^H[n]\}$ is the spatial covariance matrix for the full array. We estimate $\R$ by computing a sample covariance matrix $\eR$. A good estimate requires some degree of \emph{spatial averaging} to avoid signal cancellation, \emph{temporal averaging} to maintain true speckle statistics, and \emph{diagonal loading} to improve robustness to parameter errors ~\cite{Synnevag2009a}. Combined, these techniques will also ensure a numerically well conditioned $\eR$.

### 2019-09-30 ThesisJo/abstract.tex
2019-09-30

\input{variables}\ifMonolithic\else\input{header}\input{subheader}\fi
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\chapter{Abstract}

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\ifMonolithic\else\input{subfooter}\fi

>>>>

% \input{variables}\ifMonolithic\else\input{header}\input{subheader}\fi
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

The world is highly dynamic, constantly changing from moment to moment. An autonomous device aspiring to navigate it must be situationally perceptive and reactive. This requires numerous sensors gathering high quality data at high throughput, adaptive signal processing algorithms and powerful parallel hardware capable of near-realtime responses.

This thesis explores ways to accelerate adaptive acoustic image reconstruction and synthesis. The reconstruction process relies heavily on beamforming: the art of delaying and weighting sensor arrays to achieve desired spatial responses. A beamformer can be viewed as a spatial filter or model, and works best---as any model---when it is neither too simple nor too complex.

All models have a sweet spot in complexity, a balance point where its estimates offer an ideal compromise between bias and variance.

Such a spatial filter, or model, . model complexity should be just right. not too simple, not too complex. bias vs variance.

sensory faculty

as adaptive as possible with the least possible computational load.

gather information
dynamically analyze it
and react to it
Får

adapt
persist
prevail

Adaptive algorithm and the computation power to drive it.

Necessitates an algorithm that adapts to changing data.

Cognition is the mental action or process of acquiring knowledge and understanding through thought, experience and the senses.

The intelligence and autonomy of acoustical imaging devices require more sensory data, algorithmic adaptivity and processing power. All are vital and equally important.

improves with more high quality sensory data, sophisticated adaptive signal processing and the hardware resources to deal with it all.

more sensory data, algorithmic adaptivity and processing power.

adaptive processing and processing power.

benefits from more sensors, sensory data,   detectors rely on

autonomy improves with data amount, adaptive processing, processing power.

fast dynamic information gathering and processing.

sensory input, sophisticated information and the processing power to drive it all.

rely on fast adaptive data processing. have an intrinsic need for adaptive data processing and hardware acceleration.

Accelerated adaptive image reconstruction and synthesis.

information gathering and processing.

One use case autonomy

Key challenges that are addressed include active sonar and

- Active sonar and clinical ultrasound
- Different ways to achieve adaptivity. Analytical assessment of the data, computing weights. Approximating it. Trial-and-error. Different means to the same adaptivity
- Computational complexity
- Hardware acceleration
-

Spatial sensitivity is like a balloon; squeezing it on one side inflates it on the other.

World is highly dynamic, constantly changing from moment to moment. Necessitates an algorithm that adapts to changing data.

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% \ifMonolithic\else\input{subfooter}\fi
